---
title: "p8105_hw1_lz2949"
author: "Longyi Zhao"
date: "2023-09-17"
output: html_document
---

## Problem 1
```{r}
# installed the 'moderndive' package in console
# load the 'moderndive' library
library(moderndive)
data("early_january_weather")

```

Variables names of this data are: `r names(early_january_weather)`. There are total `r length(names(early_january_weather))` variables in the data. The size of this data is characterized by its rows and columns. This data contains `r nrow(early_january_weather)` rows and `r ncol(early_january_weather)` columns. Among these variables, temperature has a mean value of `r round(mean(early_january_weather$temp), digits=2)` degrees. 

Next step is to make a scatterplot of temperature vs. time hour and describe patterns in this plot. 

```{r}
library(tidyverse)
# plot temp vs. time_hour using ggplot 
# color points using the humid variable
ggplot(early_january_weather, aes(y=temp,x=time_hour, color=humid))+geom_point()
# save the scatterplot using ggsave function
ggsave("scatterplot_hw1_p1.pdf")
```

The temperature fluctuates in the first half of January; the overall trend was increasing. Humidity data show similar trends as the temperature did. 

## Problem 2

```{r}
# create a data frame 
# a random sample of size 10 from a standard Normal distribution
dataframe = tibble(
  random_sample = rnorm(10),
  logical_vec = random_sample>0,
  char_vec = c("I","have","a","pet","cat","called","jumy","he","is   in","black-and-white"),
  vec_factor =factor(c("level1","level2","level3","level1","level2","level3","level1","level2","level3","level3")))

# calculate the mean value of each variable 
library(tidyverse)
out_random_sample = pull(dataframe,random_sample)
out_logical_vector = pull(dataframe,logical_vec)
out_character_vector = pull(dataframe,char_vec)
out_factor_vector = pull(dataframe,vec_factor)

```

The mean value of each variable is as follows: 

1. Mean value of random sample is `r round(mean(out_random_sample),digits=4)`. 

2. Mean value of logical vector is `r mean(out_logical_vector)`. 

3. Mean value of character vector is `r mean(out_character_vector)`. 

4. Mean value of factor vector is `r mean(out_factor_vector)`. 

From the results we can see that means of character vector and factor vector variables cannot be calculated. 

# problem 3 


```{r}
# convert variables from one type to another using as.numeric function
r_s = as.numeric(out_random_sample)
r_s = as.numeric(out_logical_vector)
r_s = as.numeric(out_character_vector)
r_s = as.numeric(out_factor_vector)
```
From the result, character vector is successfully converted to numerical values while factor vector is not. 